Dynamic strategies for yaw and induction control of wind farms based on large-eddy simulation and optimization

### 2018-01-02 reply_to_reviewer2.tex

`` 
$\dots$ has a \revision{total dimension of $L_x \times L_y \times L_z = 4.8 \times 2.4 \times 1 $ km$^3$, with $L_x, L_y$, and $L_z$ the streamwise, transversal, and vertical extent of the domain respectively. This domain is discretized on a grid of $N_x \times N_y \times N_z = 256 \times 128 \times 128$ gridpoints.} The final $\dots$
''

\hrulefill

\paragraph{6. Reviewer} \textit{Line 217: Define Zh}

\paragraph{Response} Done. The manuscript is updated as follows:

``
$\dots$, \revision{and the vertical location of the turbine hubs $z_h$ is placed at the vertical mid-plane of the domain, i.e. $z_h/L_z = 0.5$,} in order to minimize $\dots$
''

@@

\paragraph{Response}

\hrulefill

\paragraph{15. Reviewer} \textit{Line 424: "...steady uniform inflow cases", would it be better to use word "constant uniform inflow cases", to avoid any confusion with steady-state analysis.}

\paragraph{Response} We have updated the manuscript to consistently use the word 'constant' instead of 'steady' when mentioning the inflow conditions. (see revised manuscript)

### 2018-01-02 MuntersMeyers_DynamicStrategiesControl_revision_WM01.tex
R2C16
R2C15

Two different sets of simulations are considered. The first set is subjected to a steady and uniform inflow with $U_\infty = 8$ m s$^{-1}$. In this case, the top and bottom boundaries are treated with a symmetry condition, \revision{and the vertical location of the turbine hubs $z_h$ is placed at the vertical mid-plane of the domain, i.e. $z_h/L_z = 0.5$,} in order to minimize any influence of vertical domain boundaries. The second set applies unsteady turbulent inflow conditions, which are generated in a pressure-driven precursor simulation of an unperturbed turbulent boundary layer with shifted periodic boundary conditions \cite{munters2016shifted}, driven by a pressure gradient of $\partial_x p_\infty/\rho = 2.5 \times 10^{-4}$ m s$^{-2}$, resulting in a freestream wind speed at hub height slightly higher than 8 m s$^{-1}$. In this second case, the top boundary is treated with a stress-free condition, and a high-Reynolds number wall model is used at the bottom. The turbines are placed at a height of $z_h/L_z = 0.1$ from the wall. The wall model adds a wall stress consistent with a rough-wall logarithmic boundary layer profile with roughness length $z_0/L_z = 10^{-4}$, resulting in a inflow turbulence intensity of about 8\% at hub height.

@@

\begin{figure}
	\includegraphics[width=\textwidth]{figure3}
	\caption{Planview at hub height $z_h$ of initial conditions for optimal control cases. \emph{a)} Uniform and steady inflow. \emph{b)} Turbulent inflow. Colors represent instantaneous streamwise velocity $\widetilde{u}_x$ in m s$^{-1}$. The black lines represent wind turbine locations. The black dashed line indicates the start of the fringe region. R1 -- R4 and C1 -- C4 indicate the numbering convention used for turbine rows and columns respectively. \label{fig:initial_conditions_flow_yaw}}
\end{figure}

@@

The current section showed results for the steady uniform inflow cases. All controlled cases increase wind-farm efficiency by a significant amount, with the yaw-enabled cases significantly outperforming cases based on induction control only. Given the current quiescent ambient flow conditions, simple yaw behavior was identified, i.e. dynamic yaw triggering wake meandering on one hand, and steady yaw resulting in wake redirection on the other hand. Moreover, the simplified control cases Ystat and Ymndr derived from this behavior attain similar wind-farm efficiencies as the optimal control cases. In the next section, unsteady and turbulent inflow conditions are applied, and it is investigated whether similar simplified control methods can be applied.

@@

Further results show that, for the current setup, all yaw-enabled cases consistently outperform cases based solely on axial induction control. However, the gap between both is significantly larger for uniform inflow than for turbulent inflow. The main reason for this is that the yawing cases perform better in the absence of inflow turbulence due to the fact that a steady flow allows wakes to be redirected more easily.

@@

\appendixtitles{yes} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixsections{multiple}
\appendix
\section{Derivation and verification of the adjoint yaw angle equation the adjoint gradient}\label{sec:app_adj_der}

>>>>

Two different sets of simulations are considered. The first set is subjected to a \revision{constant} and uniform inflow with $U_\infty = 8$ m s$^{-1}$. In this case, the top and bottom boundaries are treated with a symmetry condition, \revision{and the vertical location of the turbine hubs $z_h$ is placed at the vertical mid-plane of the domain, i.e. $z_h/L_z = 0.5$,} in order to minimize any influence of vertical domain boundaries. The second set applies unsteady turbulent inflow conditions, which are generated in a pressure-driven precursor simulation of an unperturbed turbulent boundary layer with shifted periodic boundary conditions \cite{munters2016shifted}, driven by a pressure gradient of $\partial_x p_\infty/\rho = 2.5 \times 10^{-4}$ m s$^{-2}$, resulting in a freestream wind speed at hub height slightly higher than 8 m s$^{-1}$. In this second case, the top boundary is treated with a stress-free condition, and a high-Reynolds number wall model is used at the bottom. The turbines are placed at a height of $z_h/L_z = 0.1$ from the wall. The wall model adds a wall stress consistent with a rough-wall logarithmic boundary layer profile with roughness length $z_0/L_z = 10^{-4}$, resulting in a inflow turbulence intensity of about 8\% at hub height.

@@

\begin{figure}
	\includegraphics[width=\textwidth]{figure3}
	\caption{Planview at hub height $z_h$ of initial conditions for optimal control cases. \emph{a)} Uniform and \revision{constant} inflow. \emph{b)} Turbulent inflow. Colors represent instantaneous streamwise velocity $\widetilde{u}_x$ in m s$^{-1}$. The black lines represent wind turbine locations. The black dashed line indicates the start of the fringe region. R1 -- R4 and C1 -- C4 indicate the numbering convention used for turbine rows and columns respectively. \label{fig:initial_conditions_flow_yaw}}
\end{figure}

@@

The current section showed results for the \revision{constant} uniform inflow cases. All controlled cases increase wind-farm efficiency by a significant amount, with the yaw-enabled cases significantly outperforming cases based on induction control only. Given the current quiescent ambient flow conditions, simple yaw behavior was identified, i.e. dynamic yaw triggering wake meandering on one hand, and steady yaw resulting in wake redirection on the other hand. Moreover, the simplified control cases Ystat and Ymndr derived from this behavior attain similar wind-farm efficiencies as the optimal control cases. In the next section, unsteady and turbulent inflow conditions are applied, and it is investigated whether similar simplified control methods can be applied.

@@

Further results show that, for the current setup, all yaw-enabled cases consistently outperform cases based solely on axial induction control. However, the gap between both is significantly larger for uniform inflow than for turbulent inflow. The main reason for this is that the yawing cases perform better in the absence of inflow turbulence due to the fact that a \revision{constant incoming} flow allows wakes to be redirected more easily.

@@

\appendixtitles{yes} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixsections{multiple}
\appendix
\section{Derivation and verification of the adjoint yaw angle equation and the adjoint gradient}\label{sec:app_adj_der}
